
---
title: 'Machine Learning. Peer-Graded Assignment: Prediction Assignment Writeup'
author: "Santiago Peña"
date: "18 de marzo de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(rattle)
library(corrplot)
library(knitr)
```

## Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. 

The goal of this report is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har)

The goal of this project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. 

## Loading data

The training set and data set are downloaded from urls included:

[Trainig set](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

[Testing set](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)


```{r pressure, echo=FALSE, cache=TRUE}
url1<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
training<-read.csv(url1)
url2<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
testing<-read.csv(url2)
# str(training)
# str(testing)
#Remove only if needed due to memory limitations
rm(url1,url2)
```

## Data slicing and cleaning. 

Since the so called "testing set" will be used as a validation set to check the final model, depending on the model used it could be requiered so split the training set into a testing set and a training set for cross-validation. 
That could be done by splitting the training set into a testing set and a training set partition of 25% and 75% respectively with the **createDataPartition** command. 
Crossvalidation will be done with the **TrainControl** function using a 4-fold validation method, so no further splitting is performed in the training set here.

Once the data set is splitted into a training and testing set, in order to pick the variables that correctly explain the model we firsty remove those variables that have a negligible variability in them with the **nearZeroVar** command. 
The features of variables with low variability are:


```{r, cache=TRUE}
names(training[,nearZeroVar(training)])
training2<-training[,-nearZeroVar(training)]
testing2<-testing[,-nearZeroVar(training)]
#Remove only if needed due to memory limitations
rm(training,testing)
```

Many other variables are useless since they have NA values. Those values are also removed from the test

```{r, cache=TRUE}
RemNA<-sapply(training2, function (x) any(is.na(x))) 
training3<-training2[, -which(RemNA)] 
testing3<-testing2[,-which(RemNA)]
#Remove only if needed due to memory limitations
rm(RemNA,training2,testing2)
```

Some other identification variables (first 5 columns) are also removed

```{r, cache=TRUE}
training3<-training3[,-c(1:6)]
testing3<-testing3[,-c(1:6)]
```


## Processing and Modelling

### Random forests

Within the random forests models, multiple models are developed, and the results are aggregated to improve classification rates. The algorithm involves sampling cases and variables to create a large number of deccision trees. It is a very accurate model, although it is computanionally costly.

To cross validate the model, it has been used the **TrainControl** function with a 4 fold validation and allowing parallel processing to improving computational efficiency. Unfortunately I had to reduce the training set size due to my computer processing limitations. 

```{r, cache=TRUE}
set.seed(31415)
ex<-trainControl(method="cv",number=3,allowParallel=TRUE,verboseIter=TRUE)
n<-sample(1:dim(training3)[1],11000,replace = F)
muestra<-training3[n,]
modFitRF<-train(classe ~ ., method="rf", data=muestra, prox=TRUE,trControl=ex) 
modFitRF$finalModel 
``` 

showing a very good approach approach:

```{r, cache=TRUE}
Prediction<-predict(modFitRF,training3)
table(Prediction,training3$classe)


```

It can be checked also with the remaining sample:
```{r}

Borrar<-training3[-n,]
Aver<-predict(modFitRF,Borrar)
table(Aver,Borrar$classe)

confusionMatrix(Aver,Borrar$classe) 
```




## SUMMARY AND RESULTS

From the model above, the resulting prediction results as follows:

```{r, cache=TRUE}
Borrar<-predict(modFitRF, newdata = testing3) 
Number_Files = length(Borrar)
    for (i in 1:Number_Files){
        filename =  paste0("problem_id",i,".txt")
        write.table(Borrar[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
    Borrar
```




